### 2018-11-28 latex/methods.tex
Minor changes fixing some special chars

\subsection{Molecular dynamics simulations}
We performed all-atom molecular dynamics simulations with explicit solvent for NamPT (PDB Code: 2H3D \cite{Wang2006}) with and without loop insertion. We used the AMBER99SB-ILDN force field \cite{Lindorff-Larsen2010} with the TIP3P water model \cite{Jorgensen1983} in GROMACS 5.1.2 \cite{Abraham2015}. The structures were simulated each in a box of water with distance between the solute and the box set to 0.2\,nm at a temperature of 300\,K for a total time of 1\,\textmu s. A time step of 2\,fs and the stochastic dynamics integrator were used. For the evaluation of the root mean square fluctuations (RMSF) the first 100\,ns of the simulations were omitted.

>>>>

\subsection{Molecular dynamics simulations}
We performed all-atom molecular dynamics simulations with explicit solvent for NamPT (PDB Code: 2H3D \cite{Wang2006}) with and without loop insertion. We used the AMBER99SB-ILDN force field \cite{Lindorff-Larsen2010} with the TIP3P water model \cite{Jorgensen1983} in GROMACS 5.1.2 \cite{Abraham2015}. The structures were simulated each in a box of water with distance between the solute and the box set to 0.2\,nm at a temperature of 300\,K for a total time of 1\,$\mu$s. A time step of 2\,fs and the stochastic dynamics integrator were used. For the evaluation of the root mean square fluctuations (RMSF) the first 100\,ns of the simulations were omitted.

### 2018-11-28 latex/results.tex
Replaced remaining NamPRT with NamPT in text
Minor changes fixing some special chars

To understand the functional roles and potential interplay between the three known enzymes that use Nam as substrate (NamPT, NADA and NNMT), we first conducted a comprehensive analysis of the phylogenetic distribution of these three enzymes.  As shown in Figure~\ref{fig:phylo_distribution}A, bacteria, fungi, and plants predominantly possess NADA and only a few of them harbour NamPRT. In contrast, Metazoa predominantly lost NADA and encode for NamPT together with NNMT. NADA and NamPT, the  two enzymes that initiate the two different NAD salvage pathways,  show a scattered distribution in bacteria. Co-occurrence of these enzymes is rather rare, and has mainly been found in bacteria \cite{Gazzaniga2009} and some marine invertebrates \cite{Gossmann2012FEBS}. 
NNMT seems to have arisen \textit{de~novo} or diverged rapidly in the most recent common ancestor of Ecdysozoa and Lophotrochozoa (fig.~\ref{fig:phylo_distribution}B). We were unable to find any \textit{NNMT} gene with an e-value below 0.1 in fungi or plants.

@@

Figures~\ref{fig:NamPT_affinity_Nam}E and~F visualise a direct comparison of simulations assuming different affinities of NamPT for Nam, in the presence or absence of NNMT. Interestingly, at an affinity of $K_{M}$ =~1\,$\mu$M, which is in the range of the $K_{M}$ of NADA for Nam, NAD consumption flux is only higher with NNMT present when cell division rates are low (fig.~\ref{fig:NamPT_affinity_Nam}E). If the affinity of NamPT is high enough, consumption rates are always higher with NNMT than without it. The NAD concentration is always lower with NNMT (fig.~\ref{fig:NamPT_affinity_Nam}F).

@@

The second hypothesis was  based on our model simulations that predict that NNMT might have exerted evolutionary pressure on  NamPT and that therefore the sequence insertion might have an effect on substrate binding of NamPT.  To analyse this possibility, we expressed and purified  both the wildtype and the mutant enzyme in \textit{E.~coli}., N-terminally fused to a 6xHis-tag. The size exclusion chromatography profile showed that both wildtype and mutant protein were expressed as dimers (see suppl. fig.~S3), indicating that the  mutant protein is likely to be folded correctly. The enzymatic activity was measured by NMR spectroscopy using the detection of NMN produced in the absence of ATP. Upon incubation with the NamPT inhibitor FK866 \cite{Hasmann2003} for 30 minutes, both wildtype and mutant NamPT did not produce any NMN, suggesting that binding of FK866 is not affected by the mutation (see suppl. fig.~S4). The mutant enzyme showed no detectable activity with 100 $\mu$M  Nam after 10 min of incubation at 30\textdegree{}C. With 1mM Nam enzymatic activity could be detected with the mutant enzyme. But the measured activity is reduced to 80\% of that observed for the wildtype enzyme under the same condition (fig.~\ref{fig:unresolved_loop}D). These observations suggest that the mutant enzyme is catalytically active, retains its dimeric state and sensitivity to FK866, but, has a  lower activity and affinity to Nam, supporting the predictions derived from the metabolic modelling approach.

To see whether we can find a mechanistic explanation for the reduced affinity of the mutant enzyme, we analysed different available protein structures of NamPT and tested whether  the loop insertion could potentially lead to dynamic structural rearrangements. To this end we applied homology modelling (fig.~\ref{fig:unresolved_loop}C) and molecular dynamics simulations for structures with and without the loop insertion (fig.~\ref{fig:unresolved_loop}F) Taken together we did not observe substantial structural rearrangements and the molecular dynamics simulations showed only limited structural changes upon loop insertion and we observed a mostly structurally stable catalytic core. This might be based on the fact that  all available protein structures of NamPT differ very little even at the catalytic side ( between 0.33Å and 0.95Å see supplementary table S3). Some residues close to the catalytic, nevertheless, showed slightly elevated mobilities in the wildtype structure, but, these elevated mobilities were dominated by rare events during the simulation time 1$~\mu s$. Thus, they do not appear as a robust change of structural dynamics upon loop insertion. They could nevertheless provide a hind towards reduced mobility of the mutant enzyme that explain the observed changes in affinity and activity upon loop deletion.

>>>>

To understand the functional roles and potential interplay between the three known enzymes that use Nam as substrate (NamPT, NADA and NNMT), we first conducted a comprehensive analysis of the phylogenetic distribution of these three enzymes.  As shown in Figure~\ref{fig:phylo_distribution}A, bacteria, fungi, and plants predominantly possess NADA and only a few of them harbour NamPT. In contrast, Metazoa predominantly lost NADA and encode for NamPT together with NNMT. NADA and NamPT, the  two enzymes that initiate the two different NAD salvage pathways,  show a scattered distribution in bacteria. Co-occurrence of these enzymes is rather rare, and has mainly been found in bacteria \cite{Gazzaniga2009} and some marine invertebrates \cite{Gossmann2012FEBS}.
NNMT seems to have arisen \textit{de~novo} or diverged rapidly in the most recent common ancestor of Ecdysozoa and Lophotrochozoa (fig.~\ref{fig:phylo_distribution}B). We were unable to find any \textit{NNMT} gene with an e-value below 0.1 in fungi or plants.

@@

Figures~\ref{fig:NamPT_affinity_Nam}E and~F visualise a direct comparison of simulations assuming different affinities of NamPT for Nam, in the presence or absence of NNMT. Interestingly, at an affinity of $K_{M}$~= 1\,$\mu$M, which is in the range of the $K_{M}$ of NADA for Nam, NAD consumption flux is only higher with NNMT present when cell division rates are low (fig.~\ref{fig:NamPT_affinity_Nam}E). If the affinity of NamPT is high enough, consumption rates are always higher with NNMT than without it. The NAD concentration is always lower with NNMT (fig.~\ref{fig:NamPT_affinity_Nam}F).

@@

The second hypothesis was  based on our model simulations that predict that NNMT might have exerted evolutionary pressure on  NamPT and that therefore the sequence insertion might have an effect on substrate binding of NamPT.  To analyse this possibility, we expressed and purified  both the wildtype and the mutant enzyme in \textit{E.~coli}., N-terminally fused to a 6xHis-tag. The size exclusion chromatography profile showed that both wildtype and mutant protein were expressed as dimers (see suppl. fig.~S3), indicating that the  mutant protein is likely to be folded correctly. The enzymatic activity was measured by NMR spectroscopy using the detection of NMN produced in the absence of ATP. Upon incubation with the NamPT inhibitor FK866 \cite{Hasmann2003} for 30 minutes, both wildtype and mutant NamPT did not produce any NMN, suggesting that binding of FK866 is not affected by the mutation (see suppl. fig.~S4). The mutant enzyme showed no detectable activity with 100\,$\mu$M  Nam after 10 min of incubation at 30\,\textdegree{}C. With 1mM Nam enzymatic activity could be detected with the mutant enzyme. But the measured activity is reduced to 80\% of that observed for the wildtype enzyme under the same condition (fig.~\ref{fig:unresolved_loop}D). These observations suggest that the mutant enzyme is catalytically active, retains its dimeric state and sensitivity to FK866, but, has a  lower activity and affinity to Nam, supporting the predictions derived from the metabolic modelling approach.

To see whether we can find a mechanistic explanation for the reduced affinity of the mutant enzyme, we analysed different available protein structures of NamPT and tested whether  the loop insertion could potentially lead to dynamic structural rearrangements. To this end we applied homology modelling (fig.~\ref{fig:unresolved_loop}C) and molecular dynamics simulations for structures with and without the loop insertion (fig.~\ref{fig:unresolved_loop}F) Taken together we did not observe substantial structural rearrangements and the molecular dynamics simulations showed only limited structural changes upon loop insertion and we observed a mostly structurally stable catalytic core. This might be based on the fact that  all available protein structures of NamPT differ very little even at the catalytic side ( between 0.33Å and 0.95Å see supplementary table S3). Some residues close to the catalytic, nevertheless, showed slightly elevated mobilities in the wildtype structure, but, these elevated mobilities were dominated by rare events during the simulation time 1\,$\mu$s. Thus, they do not appear as a robust change of structural dynamics upon loop insertion. They could nevertheless provide a hind towards reduced mobility of the mutant enzyme that explain the observed changes in affinity and activity upon loop deletion.
